---
title: "Extra Credit"
author: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---



\clearpage


## Extra Credit


```{r,include=FALSE}
library(pacman)
p_load(tictoc, tidyverse, data.table, tidymodels, yardstick, janitor, naniar, discrim, gmodels, knitr)

library(doParallel)
registerDoParallel(cores = 12)
```


```{r,include=FALSE}
tic()
data <- 
  fread("C:/Users/accepted_2007_to_2018Q4.csv", 
              nThread = 12, na.strings = "")
data
toc()

# subset the data for the years 2012-2014
club <- data %>%
  filter(str_detect(issue_d, '2012|2013|2014'))

var_df <- club %>% 
  map_df(~ sum(is.na(.))/length(.)) %>% 
  select_if(~ . > 0.5) %>%
  gather() 

var_na <- club %>% 
  map_df(~ sum(is.na(.))/length(.)) %>% 
  select_if(~ . > 0.5) %>%
  names()

var_na

club_df <- club %>%
  select(!one_of(var_na)) %>%                                # remove most missing value variables
  select(-id, -issue_d, -url, -zip_code, -policy_code, -application_type, -title, -emp_title, 
         -earliest_cr_line, -last_credit_pull_d, -verification_status, -last_pymnt_d, -addr_state, 
         -emp_length, -purpose) %>% 
  mutate_if(is.character, as.factor) %>%                   # convert character vars to factor vars
  mutate(loan_status_level = ifelse(loan_status == "Fully Paid", 1,  0)) %>%      # classification
  mutate(loan_status_level = as.factor(loan_status_level))
```


Subset the data for the years 2015.


```{r}
# subset the data for the years 2015
club_2015 <- data %>%
  filter(str_detect(issue_d, '2015'))

# club_2015
```


Check if the data set contains any duplicate records.


```{r}
# check duplicate data
get_dupes(club_2015)
```


Select the variables which are the same as those in data from year 2012-2014.


```{r}
var_2014 <- club_df %>% 
  names()

club_df_2015 <- club_2015 %>%
  mutate_if(is.character, as.factor) %>%         # convert character vars to factor vars
  mutate(loan_status_level = ifelse(loan_status == "Fully Paid", 1,  0)) %>%   # classification
  mutate(loan_status_level = as.factor(loan_status_level)) %>%
  select(one_of(var_2014))

# head(club_df_2015)
```


Check missing values.


```{r}
# check missing value
gg_miss_var(club_df_2015, show_pct = TRUE)
```


Summary of classification for the variable *loan_status*.


```{r}
# check loan status level
club_df_2015 %>%
  group_by(loan_status, loan_status_level) %>%
  tally()
```


```{r}
CrossTable(club_df_2015$loan_status_level, prop.chisq = FALSE)
```


Sample 10% of data set for testing 


```{r}
# sample 10% of data set
set.seed(999)
club_samp_2015 <- club_df_2015 %>% 
  slice_sample(n = 0.1*nrow(club_2015))
```


Summary of classification for the variable *loan_status*.


```{r}
#summarize the y-variable
CrossTable(club_samp_2015$loan_status_level, prop.chisq = FALSE)
```


Because there are different levels in *home_ownership* between 2 data sets, we need to drop levels in order to make 2 data sets correspond.


```{r}
levels(club_df$home_ownership)
levels(club_samp_2015$home_ownership)
```


```{r}
club_re <- club_df %>%
  mutate(home_ownership = as.character(home_ownership)) %>%
  filter(home_ownership == "ANY" | home_ownership == "MORTGAGE" | 
         home_ownership == "OWN" | home_ownership == "RENT") %>%
  mutate_if(is.character, as.factor)

levels(club_re$home_ownership)
```


Create Training nad Testing data set.


```{r}
set.seed(999)
club_samp <- club_re %>% 
  slice_sample(n = 0.1*nrow(club))
```


```{r}
set.seed(999)
samp_split <- club_samp %>% 
  initial_split(prop = 0.75)


samp_recipe <- training(samp_split) %>% 
  recipe(loan_status_level ~ .) %>% 
  step_rm(loan_status) %>%
  step_nzv(all_predictors()) %>%
  step_knnimpute(all_predictors()) %>%
  prep()

summary(samp_recipe)

tidy(samp_recipe)


samp_testing_2015 <- samp_recipe %>%
  bake(club_samp_2015)

samp_testing_2015


samp_training <- juice(samp_recipe)

samp_training
```





### Model: XGBoost


#### Setup the Model


```{r,message=FALSE, warning=FALSE}
mod_xgb_2015 <- boost_tree(trees = 300) %>% 
  set_engine("xgboost") %>%
  set_mode("classification") %>%
  fit(loan_status_level ~ ., data = samp_training)

# mod_xgb_2015
```


#### Evaluating Model Performance 


```{r out.width="85%",fig.align="center"}
# plot
plot_xgb_2015 <- mod_xgb_2015 %>% 
  predict(samp_testing_2015, type = "prob") %>% 
  bind_cols(samp_testing_2015)

plot_xgb_2015 %>% 
  roc_curve(loan_status_level, .pred_0) %>% 
  autoplot()

plot_xgb_2015 %>% 
  ggplot() + geom_density(aes(x = .pred_1, fill = loan_status_level), alpha = 0.5)
```



```{r}
# prediction on testing data set
pred_xgb_2015 <- mod_xgb_2015 %>%
  predict(samp_testing_2015) %>%
  bind_cols(samp_testing_2015)

pred_xgb_2015 %>% 
  conf_mat(truth = loan_status_level, estimate = .pred_class)

pred_xgb_2015 %>% 
  metrics(truth = loan_status_level, estimate = .pred_class)


plot_xgb_2015 %>% 
  roc_auc(loan_status_level, .pred_0)
```





